### 2013-02-04 doc/xps2013/proposal.tex
for luis

Our preliminary work on Grappa shows that it runs several graph-crunching applications (classic examples of irregular behavior) very efficiently on a commodity cluster. Our yardstick for comparison is the XMT hardware itself. Using the same number of network interfaces, Grappa is in the same ballpark as the XMT: For unbalanced tree search, Grappa is over 3X faster and shows greatly improved scalability; conversely, for breadth first search and betweenness centrality Grappa is 2.5X slower.  In Section~\ref{sec:evaluation} we explore the factors that underpin this performance. Most importantly, however, for significantly less real world cost, users can \emph{add\/} significantly more processors to a commodity cluster than an XMT machine and use Grappa to achieve scalable performance.

@@

The second challenge is the virtual to physical address translation on the cards requires the operating system to pin translations.  This removes paging flexibility from the operating system and allows user processes to (potentially unfairly) grab ahold of physical memory.  In the long run, the solution here likely lays at the intersection of what progress is occurring in virtual memory translation for I/O with virtual machines and RDMA.  Our goal will be carefully study the I/O virtualization proposals put forth by major processor manufactures and devise solutions that enable RDMA to capitalize on these developments.  \comment{weak?  -M}

The final challenge is that atomic operations via RDMA do not interact atomically with processors in a system.  Semantically it is not entire clear what happens if a processor issues an atomic operation on a memory address at the same time the network card is attempting to service a remote atomic access. \comment{maybe this is not true.  have to think some more about PCIe an what happens here.  might not be a problem.  -M}

>>>>

In this proposal we describe a body of research that extends the Grappa~\cite{grappa} runtime system.  Grappa is a runtime system that compilers target irregular applications to.  It provides several facilities: a global shared memory abstraction, user-level threading support, and communication and synchronization primitives.  Applications targeted for Grappa are not expected to have \emph{any} locality (although what locality that exists is favorably exploited), but are expected to provide many thousands of concurrent threads.

Our preliminary work on Grappa shows that it runs several graph-crunching applications (classic examples of irregular behavior) very efficiently on a commodity cluster.  For unbalanced tree search, Grappa is over 3X faster and shows greatly improved scalability compared to the Cray XMT; conversely, for breadth first search Grappa is 2.5X slower.  Most importantly, however, for significantly less real world cost, users can \emph{add} significantly more processors to a commodity cluster than an XMT machine and use Grappa to achieve scalable performance.

Grappa work is still in its infancy, however.  This research proposal seeks funding to extend Grappa in the following directions:

\begin{itemize}
\item Presently, applications are hand-coded to the Grappa interface.  A significant research thrust of the proposed work is to ease this programming burden by language and compilation support.

\item Our preliminary work has also revealed limitations of the current Grappa runtime system in the area of network scaling.  This proposal seeks funding to enhance the network layer and explore the architecture/software interactions of remote direct memory access (RDMA).

\item The size of data being processed by large systems today (hundreds of terabytes) dwarfs what can fit into a single system. Obviously multiple system nodes are required to maintain and access all of this data efficiently.  However, even with hundreds of channels to persistent storage, loading this data into DRAM can prove a bottleneck.  The Grappa runtime is designed to make applications tolerate extremely high memory latency, thus an idea we wish to explore is using solid state disks as main memory.

\item Presently there does not exist a comprehensive benchmark suite of irregular and graph applications.  We also seeking funding through this proposal to put together just such a

\end{itemize}

%% Roadmap

@@

The second challenge is the virtual to physical address translation on the cards requires the operating system to pin translations.  This removes paging flexibility from the operating system and allows user processes to (potentially unfairly) grab ahold of physical memory.  In the long run, the solution here likely lays at the intersection of what progress is occurring in virtual memory translation for I/O with virtual machines and RDMA.  Our goal will be to carefully study the I/O virtualization proposals put forth by major processor manufactures and devise solutions that enable RDMA to capitalize on these developments.  \comment{weak?  -M}

The final challenge is that atomic operations via RDMA do not interact atomically with processors in a system.  Semantically it is not entirely clear what happens if a processor issues an atomic operation on a memory address at the same time the network card is attempting to service a remote atomic access. \comment{maybe this is not true.  have to think some more about PCIe an what happens here.  might not be a problem.  -M}
